Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-01-13 paper-1/manuscript.tex
Rev sec Results:DD (F2) for notation consistency

We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs. Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the \dev{} sequence to A in the \std{} sequence). In other words, we used the \msc{} sequence as a neutral baseline where the target stimulus was presented no more often than in the \dev{} sequence, but no expectation of a competing stimulus B could be formed. To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, denoted as $R^A_{dev}$ and $R^A_{msc}$, defining a deviance detection index (DDI) as
\begin{equation}
    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}} \label{eqn-ddi}
\end{equation}

\begin{figure*}%
    \centering
    \includegraphics{fig2.pdf}
    \caption{}
    \label{fig2}
\end{figure*}
\textbf{Fig. 2} Deviance detection in our model. \textbf{a} Sample network target response intensity by trial type. Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step. Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section). \textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence. The underlying data are the same as in panel a. Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
\textbf{c} Response sizes $R^A_{dev}$ and $R^A_{std}$, normalized by $R^A_{msc}$, across networks and stimuli (n = 120). In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
\textbf{d} Deviance detection indices (DDI, see Equation {eqn-ddi}) across networks and stimuli

>>>>

We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs. Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the \dev{} sequence to A in the \std{} sequence). In other words, we used the \msc{} sequence as a neutral baseline where the target stimulus was presented no more often than in the \dev{} sequence, but no expectation of a competing stimulus B could be formed. To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, defining a deviance detection index (DDI) as
\begin{equation}
    DDI = \frac{\mean{\R[dev]{A}{}} - \mean{\R[msc]{A}{}}}{\mean{\R[dev]{A}{}} + \mean{\R[msc]{A}{}}} \label{eqn-ddi}
\end{equation}

\begin{figure*}%
    \centering
    \includegraphics{fig2.pdf}
    \caption{}
    \label{fig2}
\end{figure*}
\textbf{Fig. 2} Deviance detection in our model. \textbf{a} Sample network target response intensity by trial type. Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step. Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section). \textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence. The underlying data are the same as in panel a. Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
\textbf{c} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120). In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
\textbf{d} Deviance detection indices (DDI, see Equation {eqn-ddi}) across networks and stimuli
